The Inner Mass-Ratio Distribution of Intermediate-Mass Stars

### 2016-02-11 paper/ms.tex
Fixed MRD figure.
Added some model comparison to discussion section.

\begin{align}
 P_1(q|\vec{\theta}) &= \begin{cases}
      \hfill \theta_1 \hfill & q \in \Delta_1 \\
      \hfill \theta_2 \hfill & q \in \Delta_2 \\
      \hfill \ldots \\
      \hfill \theta_7 \hfill & q \in \Delta_7
     \end{cases} \label{eqn:P1} \\
 P_2(q|\vec{\theta}) &= \frac{A}{\sqrt{2\pi\sigma^2}} e^{-\frac{(q - \mu)^2}{2\sigma^2}} \label{eqn:P2}
\end{align}
The constant $A$ in the gaussian distribution is a renormalization factor such that the distribution is only defined from $0 < q < 1$:

@@

We fit both distributions via Importance Nested Sampling with the MultiNest code \citep{multinest}. Following \citet{Foreman2014}, we apply a smoothing prior on the parameters $\vec{\theta}$ for the histogram model:

@@

@@

We already know $P(q|\vec{\theta})$ (Equations \ref{eqn:P1} and \ref{eqn:P2}). We estimate $P(\mathrm{obs}|q)$ by simulating a very large sample of binary stars:

@@

We show the resulting fits in Figure \ref{fig:mrd}. The $1\sigma$ uncertainties in the bin heights from the histogram model are shown as error bars, and we overplot 500 samples of the gaussian fit to show the spread allowed by the data.

\begin{figure}
\includegraphics[width=\columnwidth]{MRD_normal.pdf}
\caption{Mass ratio distribution for our sample. The data was fit to both a histogram and a gaussian distribution. The histogram is shown in the solid blue blocks, with $1 \sigma$ uncertainties marked with error bars. The variance of the gaussian fit is shown with 500 samples from the posterior probability distribution for the parameters.}
\label{fig:mrd}
\end{figure}

@@

The most striking feature of the mass ratio distribution shown in Figure \ref{fig:mrd} is the turnover at intermediate q, which the gaussian fit finds occurs at $q = 0.36 \pm 0.04$ and which excludes a power law fit (including a flat distribution).

TODO: Compare to chabrier IMF, power law, and flat distribution quantitatively. Bayesian evidence ratios?

@@

%              &                &                &                &          MJD &            &  exptime &     snr \\
%         star &            SpT &             ra &            dec & (JD-2450000) & instrument &  (sec)   &

>>>>

\begin{align}
 P_1(q|\vec{\theta}) &= \begin{cases}
      \hfill \theta_1 \hfill & q \in \Delta_1 \\
      \hfill \theta_2 \hfill & q \in \Delta_2 \\
      \hfill \ldots \\
      \hfill \theta_7 \hfill & q \in \Delta_7
     \end{cases} \label{eqn:P1} \\
 P_2(q|\vec{\theta}) &= \frac{A}{\sqrt{2\pi\sigma^2}} e^{-\frac{(q - \mu)^2}{2\sigma^2}} \label{eqn:P2} \\
 P_3(q|\vec{\theta}) &= (1-\gamma)q^{-\gamma} \label{eqn:P3}
\end{align}
The constant $A$ in the gaussian distribution is a renormalization factor such that the distribution is only defined from $0 < q < 1$:

@@

We fit all distributions via Importance Nested Sampling with the MultiNest code \citep{multinest}. Following \citet{Foreman2014}, we apply a smoothing prior on the parameters $\vec{\theta}$ for the histogram model:

@@

The power law fit has only one parameter ($\gamma$); we use a uniform prior in the fit.

@@

We already know $P(q|\vec{\theta})$ (Equations \ref{eqn:P1} - \ref{eqn:P3}). We estimate $P(\mathrm{obs}|q)$ by simulating a very large sample of binary stars:

@@

We show the resulting fits in Figure \ref{fig:mrd}. The $1\sigma$ uncertainties in the bin heights from the histogram model are shown as error bars, and we overplot 500 samples of the gaussian fit to show the spread allowed by the data. The best-fit power law is plotted with a green dot-dashed line, as well as the mass ratio distribution expected from random pairing from the Kroupa IMF in yellow.

TODO: Put the MCMC samples up on zenodo or something.

\begin{figure}
\includegraphics[width=\columnwidth]{MRD_total.pdf}
\caption{Mass ratio distribution for our sample. The data was fit to a histogram, a gaussian distribution, and a power law. The histogram is shown in the solid blue blocks, with $1 \sigma$ uncertainties marked with error bars. The variance of the gaussian fit is shown with 500 samples from the posterior probability distribution for the parameters. We also show the best-fit power law and the mass ratio distribution from random pairing of the Kroupa initial mass function (IMF).}
\label{fig:mrd}
\end{figure}

@@

The most striking feature of the mass ratio distribution shown in Figure \ref{fig:mrd} is the turnover at intermediate q, which the gaussian fit finds occurs at $q = 0.36 \pm 0.04$. Although the power law fit and the Kroupa IMF are visually very poor fits to the data, we formally compare the models to ensure that the different form is not an artifact of binning or simple noise. We make the comparison using the posterior odds:

\begin{equation}
\mathrm{Odds} = \frac{\int_{\vec{\theta_1}} P_1(q|\vec{\theta_1}) P(\vec{\theta_1})}{\int_{\vec{\theta_2}} P_2(q|\vec{\theta_2}  P(\vec{\theta_2})} \equiv \frac{Z_1}{Z_2}
\end{equation}

The integrals are estimated as part of the nested sampling algorithm in the MultiNest code. The odds ratio comparing the gaussian to power law fit is $Z_\mathrm{gauss} / Z_\mathrm{power} = 69 \pm 1.1 \e{7}$ indicating a very strong preference for the gaussian model. We also compare to the mass ratio distribution expected for random pairing from the Kroupa IMF. In this case, there are no free parameters so the evidence integral just becomes the likelihood function (Equation \ref{eqn:money}). The corresponding odds ratio is  $Z_\mathrm{gauss} / Z_\mathrm{kroupa} = 333 \pm 4 \e{17}$, again indicating a very strong preference for the gaussian fit.

@@
